---
title: "EdgeR_2"
author: "Dennis Amnebrink"
date: '2022-10-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r constants, message=FALSE, warning=FALSE}
SIGNIFICANCE = 0.05
DIFF_TRESHOLD = 2.5
MIN_COUNT = 15
MIN_SAMPLE = 2
```

```{r libraries}
library(data.table)
library(edgeR)
library(tidyverse)
library(kableExtra)
```

```{r read-data}
#Count file for all orfs, tpm is calculated per sample
bbmap <- fread("../data/bbmap_counts.tsv.gz", sep = "\t") %>%
              filter(count > 0)  %>%                               # removing 0 counts to reduce size of table
              mutate(Geneid = str_replace(Geneid, "[0-9]*_",""),   # Joining in chr and geneid to match format of eggnog naming.
                     orf = paste(Chr,Geneid, sep = "_"),
                     sample = str_extract(sample, "P[0-9]*_[0-9]*")) %>% # Removing extra numbers in sample to match sample_ID format
              dplyr::select(-Geneid,-Chr) %>%
              dplyr::select(orf,Start,End,Strand,Length,sample,count,tpm) 

#Reading in annotations
eggnogs <- fread("../data/eggnog_annotations.tsv.gz", sep = '\t')

#File to connect NGI ID with our own classification ID, add the µ-values to this file for testing later, or do that in a sep file?
sample_ID <- read_tsv("../data/J.Pinhassi_21_02_sample_info.txt") %>%
  dplyr::select(1:2) %>%
  dplyr::rename(sample = 1, sample_name = 2) %>%
  mutate(sample_name = str_replace(sample_name, "[0-9]*:","")) %>%
  separate(sample_name, c("treatment","timepoint"), sep = ",") %>%
  mutate(treatment = gsub("NT","TN", treatment)) %>%
  mutate(sample_name = paste(timepoint, treatment, sep = "_")) %>%
  separate(treatment, c("treatment","replicate"), sep = -1) %>%
  mutate(tre_rep = gsub("_","", sample_name)) %>%
  mutate(tre_rep = toupper(tre_rep)) %>%
  mutate(day = case_when( timepoint == "t3" ~ 10,
                          TRUE ~ 17
                          )
         ) %>%
  as.data.table()

#Reading in taxonomic annotation of reads
taxonomy <- read_tsv("../data/eukulele_phylodb.tsv") %>%
  select(-species) %>%
  rename(species = "genus", genus = "family", family = "order", order = "class", class = "phylum", phylum = "kingdom") %>%
  as.data.table()
```

```{r prok-filter, message=FALSE, warning=FALSE, cache=TRUE}
#Removing non-prokreads as well as recalculating tpm over prok frac.
bbmap_p <- bbmap %>%
  semi_join(taxonomy %>% filter(domain %in% c("Bacteria", "Archaea")), 
            by = "orf") 
```

# EdgeR test
```{r}
design <- model.matrix(~ 0 + factor(c(1,1,1,2,2,2,3,3,3,4,4,4)))
colnames(design) <- c("warm_control","cold_control","cold_nutrient","warm_nutrient")

#Creating DGElist
dgelist <- bbmap_p %>%
  group_by(orf) %>%
  filter(sum(count) >= MIN_COUNT, n() >= MIN_SAMPLE) %>%
  ungroup() %>%
  semi_join(sample_ID %>% filter(day == '10'), by = 'sample') %>%
  inner_join(eggnogs, by = 'orf') %>%
  group_by(seed_eggNOG_ortholog, sample) %>%
  summarise(count = sum(count), .groups = 'drop') %>%
  ungroup() %>%
  pivot_wider(names_from = sample, values_from = count, values_fill = 0) %>%
  as.data.frame() %>%
  tibble::column_to_rownames('seed_eggNOG_ortholog') %>%
  DGEList() %>%
  calcNormFactors() %>%
  estimateDisp(design) %>%
  glmQLFit(design, robust = TRUE)



  
  

#Creating contrasts
my.contrasts <- makeContrasts(cold_nutrientvscold_control=cold_nutrient-cold_control,
                              warm_nutrientvswarm_control=warm_nutrient-warm_control,
                              levels=design)



#Creating contrasts
qlf.NvsC <- glmQLFTest(dgelist, contrast=my.contrasts[,"cold_nutrientvscold_control"])

qlf.TNvsT <- glmQLFTest(dgelist, contrast=my.contrasts[,"warm_nutrientvswarm_control"])

#Combining tables and mutating fdr-values
qlf.NvsC$table %>%
  as.data.frame() %>%
  rownames_to_column("seed_eggNOG_ortholog") %>%
  mutate(contrast = "NvsC") %>%
  bind_rows(
    qlf.TNvsT$table %>%
    as.data.frame() %>%
    rownames_to_column("seed_eggNOG_ortholog") %>%
    mutate(contrast = "TNvsT") %>%
    as.data.frame()
  ) %>%
  mutate(fdr = p.adjust(PValue, method = "fdr")) %>%
  mutate(type = 
           case_when(
             (fdr <= SIGNIFICANCE & logFC >= DIFF_TRESHOLD) ~ "Up",
             (fdr <= SIGNIFICANCE & logFC <= -DIFF_TRESHOLD) ~ "Down",
             TRUE ~"non-DE"
           )) %>%
  mutate(type = forcats::as_factor(type) %>% fct_relevel("Up","non-DE","Down")) %>%
  write_tsv("../results/edgeR_contrasts.tsv")
```

## Sanity checking
```{r}
contrasts <- read_tsv("../results/edgeR_contrasts.tsv")

#Plotting the top20 seed_eggNOG_otrhologs to see if contrasts make sense, using tpm
plot_obj <- bbmap_p %>% 
  semi_join(sample_ID %>% filter(day == '10'), by = 'sample') %>% #Filtering out relevant samples
  inner_join(eggnogs %>% distinct(orf, seed_eggNOG_ortholog), by = "orf") %>%
  group_by(sample, seed_eggNOG_ortholog) %>% #Summarising counts per seed_eggNOG ortholog and sample
  summarise(sum_tpm = sum(tpm)) %>%
  ungroup() %>%
  inner_join(sample_ID %>% filter(day == '10'), by = 'sample') %>%
  group_by(treatment, seed_eggNOG_ortholog) %>% #Averaging counts per treatment and seed_eggNOG_ortholog
  summarise(mean_tpm = mean(sum_tpm)) %>%
  ungroup() 
```

```{r}
#Plotting the genes with the largest logfold-change
top50_NvsC <- contrasts %>% 
  filter(contrast == "NvsC" & fdr <= 0.05) %>%
  arrange(desc(logFC)) %>%
  slice(1:50) %>%
  pull(seed_eggNOG_ortholog)

plot_obj %>%
  filter(treatment %in% c("N","C")) %>%
  filter(seed_eggNOG_ortholog %in% top50_NvsC) %>%
  ggplot(mapping = aes(x = seed_eggNOG_ortholog, y = mean_tpm, fill = treatment)) +
  geom_col() +
  theme(axis.text.x = element_text(hjust = 1, angle = 60)) +
  ggtitle("NvsC top 50 logFC")
```

```{r}
#Plotting the genes with the largest negative logfold-change
top50_CvsN <- contrasts %>% 
  filter(contrast == "NvsC" & fdr <= 0.05) %>%
  arrange(logFC) %>%
  slice(1:50) %>%
  pull(seed_eggNOG_ortholog)

plot_obj %>%
  filter(treatment %in% c("N","C")) %>%
  filter(seed_eggNOG_ortholog %in% top50_CvsN) %>%
  ggplot(mapping = aes(x = seed_eggNOG_ortholog, y = mean_tpm, fill = treatment)) +
  geom_col() +
  theme(axis.text.x = element_text(hjust = 1, angle = 60)) +
  ggtitle("CvsN")
```

## Overview stats

### Table of edgeR-analysis
This table shows the number of significantly upregulated/downregulated KO-terms as well as non-significant KO-terms per contrast. 
```{r}
kable(contrasts %>%
  group_by(contrast, type) %>%
  tally() %>%
  ungroup() %>%
  spread(contrast, n, fill = 0)) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  add_header_above(c("Table 1. EdgeR stats contrasts" = 3), font_size = 18) %>%
  scroll_box(width = "100%", height = "200px")
```

### Plot of edgeR-analysis
```{r}
contrasts %>% 
  ggplot(mapping = aes(x=logCPM, y = logFC, colour=type, size = type)) +
  geom_point()  +
  scale_size_manual(values = c("Down" = 1.5, "Up"=1.5, "non-DE" = 0.3)) +
  scale_colour_manual(values = c("Down" = "Blue","Up" = "Red", "non-DE" = "Black")) +
  theme_minimal() +
  geom_hline(aes(yintercept = DIFF_TRESHOLD), linetype = "dashed", color = "green") +
  geom_hline(aes(yintercept = -DIFF_TRESHOLD), linetype = "dashed", color = "green") +
  facet_wrap(~ contrast)

ggsave("../results/edgeR_eggnoG_contrasts.pdf")
```
## Summary over changes in large COG-categories between treatments (facet wrap on contrast, and colour by treatment)
```{r COGCats_NvsC}
#Plotting the contrast for NvsC
plot_obj <- bbmap_p %>%
  semi_join(sample_ID %>% filter(day == "10"), by = "sample") %>%
  inner_join(eggnogs, by = 'orf') %>%
  group_by(seed_eggNOG_ortholog, sample) %>%
  summarise(sum_tpm = sum(tpm), .groups = 'drop') %>%
  ungroup() %>%
  inner_join(eggnogs %>% 
               distinct(seed_eggNOG_ortholog, best_og_cat), 
             by = "seed_eggNOG_ortholog") %>%
  separate_rows(best_og_cat, sep = '(?<=.)(?=.)') %>%
   inner_join(contrasts %>% 
               filter(contrast == "NvsC") %>% 
               select(seed_eggNOG_ortholog, type), 
             by = "seed_eggNOG_ortholog") %>%
  #filter(type != "non-DE") %>%
  group_by(sample, best_og_cat, type) %>%
  summarise(sum_tpm = sum(sum_tpm)) %>%
  ungroup() %>%
  inner_join(sample_ID, by = "sample") %>% 
  group_by(best_og_cat, treatment, type) %>%
  summarise(mean_tpm = mean(sum_tpm)) %>%
  ungroup()

## Fix here by arranging in descending values based on the up-regulated genes.

cat_order <- plot_obj %>%
  filter(treatment %in% c("N")) %>%
  filter(type == "Up") %>%
  mutate(best_og_cat = factor(best_og_cat) %>% forcats::fct_reorder(mean_tpm, .desc = TRUE)) %>%
  arrange(mean_tpm) %>%
  pull(best_og_cat) %>%
  as.vector()

NvsC_catvec <- cat_order
  
plot_obj %>% 
  filter(treatment %in% c("N","C")) %>%
  ggplot(mapping = aes(x = fct_relevel(best_og_cat, NvsC_catvec), y = mean_tpm, fill = fct_relevel(type, c("non-DE", "Down","Up")))) +
  geom_col() +
  theme_minimal() +
  facet_wrap(~ treatment) +
  coord_flip() +
  theme(axis.text.x = element_text(hjust = 1, angle = 60)) +
  ggtitle("COG-category expression of gene expression analysis N and C-treatment")

NvsC_cont <- last_plot()

ggsave("../results/NvsC_COG_cats.png")
```

```{r COGCats_TNvsT}
plot_obj <- bbmap_p %>%
  semi_join(sample_ID %>% filter(day == "10"), by = "sample") %>%
  inner_join(eggnogs, by = 'orf') %>%
  group_by(seed_eggNOG_ortholog, sample) %>%
  summarise(sum_tpm = sum(tpm), .groups = 'drop') %>%
  ungroup() %>%
  inner_join(eggnogs %>% 
               distinct(seed_eggNOG_ortholog, best_og_cat), 
             by = "seed_eggNOG_ortholog") %>%
  separate_rows(best_og_cat, sep = '(?<=.)(?=.)') %>%
   inner_join(contrasts %>% 
               filter(contrast == "TNvsT") %>% 
               select(seed_eggNOG_ortholog, type), 
             by = "seed_eggNOG_ortholog") %>%
  #filter(type != "non-DE") %>%
  group_by(sample, best_og_cat, type) %>%
  summarise(sum_tpm = sum(sum_tpm)) %>%
  ungroup() %>%
  inner_join(sample_ID, by = "sample") %>% 
  group_by(best_og_cat, treatment, type) %>%
  summarise(mean_tpm = mean(sum_tpm)) %>%
  ungroup()

## Fix here by arranging in descending values based on the up-regulated genes. 

cat_order <- plot_obj %>%
  filter(treatment %in% c("TN")) %>%
  filter(type == "Up") %>%
  mutate(best_og_cat = factor(best_og_cat) %>% forcats::fct_reorder(mean_tpm, .desc = TRUE)) %>%
  arrange(mean_tpm) %>%
  pull(best_og_cat) %>%
  as.vector()

TNvsT_catvec <- c("A","W",cat_order)
  
#The last two categories in the vector are missing since they do do not have any upregulated genes (A,W), so add to bottom. To add to the picture I had in mind, grey them out and do not include them in the ranking plot.
plot_obj %>% 
  filter(treatment %in% c("TN","T")) %>%
  ggplot(mapping = aes(x = fct_relevel(best_og_cat, TNvsT_catvec), y = mean_tpm, fill = fct_relevel(type, c("non-DE", "Down","Up")))) +
  geom_col() +
  theme_minimal() +
  facet_wrap(~ treatment) +
  coord_flip() +
  theme(axis.text.x = element_text(hjust = 1, angle = 60),legend.position = "bottom") +
  ggtitle("COG-category expression of gene expression analysis TN and T-treatment") +
  scale_fill_discrete(name = "Type")

TNvsT_cont <- last_plot()



ggsave("../results/TNvsT_COG.png")

NvsC
NvsC_cont
TNvsT_cont

```
Here add the lines between COG-categories between the COG-categoriew to really show the difference between treatments.
```{r contrast_comparison}

# Do a classic line chart, mutate in line numbers to arrange the categories, and use the categories as labels instead.
#This vector is what fucks up the plotting, so do this differently
TNvsT_catvec <- c("A","W",TNvsT_catvec)

NvsC_rank_df <- NvsC_catvec %>% 
  as.data.frame() %>%
  mutate(rank = row_number()) %>%
  rename(category = 1)  %>%
  mutate(contrast = "NvsC")

TNvsT_rank_df <- TNvsT_catvec %>%
  as.data.frame() %>%
  mutate(rank = row_number()) %>%
  rename(category = 1) %>%
  mutate(contrast = "TNvsT")

cat_df <- NvsC_rank_df %>%
  bind_rows(TNvsT_rank_df)


mutate(best_og_cat = factor(best_og_cat) %>% forcats::fct_reorder(mean_tpm, .desc = TRUE))

#Trying ewith cats straight
cat_df %>%
  mutate(category = as.factor(category) %>% forcats::fct_relevel(NvsC_catvec)) %>%
  ggplot(mapping = aes(x = contrast, y = category, group = rank)) +
  geom_line() +
  scale_y_discrete()

#Trying with ranks instead an adding labels later
cont_order <- cat_df %>%
  mutate(crank = as.integer(rank)) %>%
  mutate(new_contrast = as.integer(ifelse(contrast == "NvsC", 1,2))) %>%
  ggplot(mapping = aes(x = new_contrast, y = rank, group = category)) +
  geom_line() +
  scale_y_continuous(breaks = unique(cat_df$rank), labels = NvsC_catvec,
                     sec.axis = sec_axis(trans=~.*1, breaks = unique(cat_df$rank) ,name="", labels = TNvsT_catvec)) +
  scale_x_continuous(breaks = c(1,2), labels = unique(cat_df$contrast)) +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.text.x = element_blank()) +
  ylab("") +
  xlab("")



trial_df <- cbind(as.data.frame(TNvsT_catvec), as.data.frame(NvsC_catvec))

#Data in wide format for parcoords

parcoords(trial_df, rownames = FALSE, reorderable = TRUE,brushMode = "1d-multi")


last_plot()
```

```{r patchwork_contrasts}

combined <- NvsC_cont  + ggtitle("") + xlab("") +
  cont_order +
  TNvsT_cont + ggtitle("") + xlab("") +
   plot_annotation(title = 'Contrast response comparison') 


 NvsC_cont  + ggtitle("") + xlab("") + theme(legend.position = "none", axis.text.y = element_blank()) +
   plot_spacer() +
  cont_order +
   plot_spacer() +
  TNvsT_cont + ggtitle("") + xlab("")  + theme(axis.text.y = element_blank()) +
  plot_annotation(title = 'Contrast response comparison') +
  plot_layout(ncol = 5, nrow = 1, widths = c(1,-0.08,0.1,-0.1,1), byrow = FALSE)
 
 
 
ggsave("../results/contrasts_COG.pdf")


  


ggsave("../results/contrast_comparison.pdf")

```

It is noteworthy that the majorty of the genes downregulated (that is more elevated in the C treatment) belong to category P. While a wider response is seen in the N-treatment.
```{r}
#Plotting the contrast for TNvsT
plot_obj <- bbmap_p %>%
  semi_join(sample_ID %>% filter(day == "10"), by = "sample") %>%
  inner_join(eggnogs, by = 'orf') %>%
  group_by(seed_eggNOG_ortholog, sample) %>%
  summarise(sum_tpm = sum(tpm), .groups = 'drop') %>%
  ungroup() %>%
  inner_join(eggnogs %>% 
               distinct(seed_eggNOG_ortholog, best_og_cat), 
             by = "seed_eggNOG_ortholog") %>%
  separate_rows(best_og_cat, sep = '(?<=.)(?=.)') %>%
   inner_join(contrasts %>% 
               filter(contrast == "TNvsT") %>% 
               select(seed_eggNOG_ortholog, type), 
             by = "seed_eggNOG_ortholog") %>%
  filter(type != "non-DE") %>%
  group_by(sample, best_og_cat, type) %>%
  summarise(sum_tpm = sum(sum_tpm)) %>%
  ungroup() %>%
  inner_join(sample_ID, by = "sample") %>% 
  group_by(best_og_cat, treatment, type) %>%
  summarise(mean_tpm = mean(sum_tpm)) %>%
  ungroup()

plot_obj %>% 
  filter(treatment %in% c("TN","T")) %>%
  ggplot(mapping = aes(x = fct_relevel(best_og_cat, cat_order), y = mean_tpm, fill = treatment)) +
  geom_col() +
  theme_minimal() +
  facet_wrap(~ type) +
  coord_flip() +
  theme(axis.text.x = element_text(hjust = 1, angle = 60)) +
  ggtitle("COG-category expression of gene expression analysis TN and T-treatment")
```
In the TN vs T treatment instead a diverse response in seen spanning the majority of categories for both treatment TN and T.
### Number of eggnogs in each cog-category per contrast
```{r}
contrasts %>%
  filter(type != "non-DE") %>%
  inner_join(eggnogs %>% 
               distinct(seed_eggNOG_ortholog, best_og_cat) %>%
               separate_rows(best_og_cat, sep = '(?<=.)(?=.)'),
             by = "seed_eggNOG_ortholog") %>%
  group_by(contrast, best_og_cat, type) %>%
  tally() %>%
  ungroup() %>%
  ggplot(mapping = aes(x = best_og_cat, y = n, fill = type)) +
  geom_col() +
  theme_minimal() +
  facet_wrap(~ contrast) +
  coord_flip()
```


